---
title: "Exploratory Analysis"
output: html_document
---

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

### Authors

Aisha Abdulkarim (aishaa19@uw.edu), Giselle Calvario (gcalva@uw.edu), 
Sumaya Mahamed (sumaha@uw.edu), Aditi Manjunath (aditim4@uw.edu) 

### Date

May 18, 2024

### Abstract

The objective of this project is to conduct an analysis of the foundation shade offerings between two highly sought and leading cosmetic brands, Fenty Beauty and Mac Cosmetics. We are collecting and analyzing our data from, *The Pudding: Beauty Brawl*. We intend to evaluate which company provides a more diverse and inclusive range of foundation shades to their clients. Furthermore, we plan to explore the reasons why one might provide more shade range compared to the other by looking more into their Saturation, Hues, and Lightness. 

### Keywords

Beauty: Foundation shades
Inclusivity: Shade Diversity 
Inclusivity: Possible clientele

### Introduction  

Our project is focused on the inclusivity of foundation ranges with two brands such as Fenty Beauty and Mac. For many years, there has been ongoing demand and concerns regarding inclusivity in the beauty industry. It is common knowledge, how diversity is glaringly insufficient and undesirable in today's progressive society. The darker shades of foundation on the color spectrum are usually limited and not many people of color have access to a wider range of shades within their skin tone. Despite people of color perpetually voicing their concerns about the lack of diversity. Unfortunately, many well-known brands disregard these claims and instead, they consistently cater to a lighter range of skin tones. No matter how many times consumers raise their voices, it has become apparent that disregarding the inclusivity of a certain population is a very detrimental oversight. Considering the fact that customers can wield their purchasing power anytime for a change. After many years, now the beauty industry is slowly evolving and some changes have been pioneered with brands like Fenty Beauty. Especially brands like Fenty Beauty a Black-owned brand, is known for advocating for inclusivity and her foundation ranges are more diverse than any other brands. 

Some research questions that we want to address in our project are what ranges of foundation shades are offered compared to different brands? Also how diverse are the lightness, saturation, and hues of each of the undertones of the brands? and how each brand shows inclusivity within its foundation ranges?

These questions are important to address because we want to investigate how prominent it is to have inclusivity when it comes to the beauty industry. Moreover, we want the beauty industry to cater to all skin tones and not be biased when it comes to representing people of color.

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

This analysis can have many implications. For example, Technologists can explore the underlying mechanisms behind both of the shade-matching algorithms between the brands. By finding out the hue, saturation, and lightness of the foundation shades, it could lead to a better understanding of color-matching algorithms and technology which could lead to more inclusivity as well as accuracy in shades in makeup products in the future. In addition, policymakers can gain insight into the implications of shade inclusivity in the beauty industry for regulations and guidelines. This can help them explore ways to enforce diversity and inclusivity standards in the cosmetic industry and manufacturing of products. This can help ensure that consumers have access to products suiting their diverse skin tones. The interesting implication that the comparison of hue, saturation, and lightness can have is, does the brand has many shades where the shades are pretty similar and not as spread out, or are the shades more spread out and different from each other? This is what we wanted to explore with this analysis. 


### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1: Histogram

```{r, echo = FALSE, code = readLines("Chart 1.R")}
```

The above data set shows the comparison of lightness shades of the two brands Fenty and Mac. Observing the data set you can see how Fenty Foundation has more darker shades than the brand Mac. On the other hand, Mac has more frequency of lightness foundation than Fenty. For example, in the dataset, Mac has six brands of foundation that appear to be lighter and one shade that is darker. Additionally, in this dataset, the brand Mac appears to cater to more lightness instead of having an equal distribution of shade ranges like Fenty. Moreover, one can make the inference that if you are a person of color then a wiser choice would be to use Fenty Foundation as it appears to be more inclusive than Mac.

### Chart 2: Box Plot

```{r, echo = FALSE, code = readLines("Chart 2.R")}
```

I used a box plot to compare the hues of foundations from Fenty and Mac. From this plot it is evident that is more inclusive of medium and darker skin tones. The range is greater for Fenty foundations hues as well.


### Chart 3: Bar Graph

This Bar Chart is included to show the difference in Saturation Levels of Foundation Shades between *Fenty Beauty and Mac Cosmetics*. According to *Iredale, J.*, saturation is defined as “How Intense the color is.”. The purpose of this chart is to show which brand had the highest saturation level and answer our question of, What role does Saturation levels play in Shade ranges. 

```{r, echo = FALSE, code = readLines("Chart 3.R")}
```

From the bar chart, “Fenty vs. Mac: Saturation Level Differences”, we are evaluating the saturation levels, from the lowest at 0.0 and the highest at 0.8. As we observe the chart, we can identify that Fenty has a higher saturation level than Mac. Why is that? When we look at the number of dark shades Fenty produces compared to Mac, Fenty is the overall winner in producing Dark shades due to its high customer base who have a darker skin tone, while also including the fact that the Company is owned by a woman of color.  Darker shades are mainly attributed to how saturated the shade is compared to lighter shades/colors. When shades have higher results of saturation it results in foundation shades or colors being darker and when the saturation is low the lighter the color, in this case, the lighter the foundation shade. 

According to our dataset, Mac had released 42 Foundation shades, while Fenty had released 39 Foundation shades when this data was taken. Even though Mac seems to have the advantage because they have released more shades compared to Fenty, why is their saturation levels low? This might have been influenced by them releasing foundation shades that were lighter and or releasing a smaller variation of dark shades than they have for lighter shades. This is not just something that relates to Mac and Fenty but the majority of beauty companies tend to drop more light-based shades in the early stages of production and then expand to darker shades as they get more consumers. 


### Citations 

Iredale, J. (n.d.). What color makeup should I wear? color theory simplified. jane iredale. https://janeiredale.com/blogs/makeup-blog/what-color-makeup-should-i-wear-color-theory-simplified#:~:text=Saturation%20is%20how%20intense%20the,have%20a%20lot%20of%20saturation.  