---
title: "Exploratory Analysis"
output: html_document
---

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

### Authors

Aisha Abdulkarim, Giselle Calvorio, Sumaya Mahamed, Aditi Manjunath  

### Date

May 17, 2024

### Abstract

No more than three sentences that summarize your project. Focus on the very most important aspects. For example: (1) "Our main question is .... This question is important because .... To address the question, we will ...." (2) "We are concerned with ..., because .... To address this concern, we plan to ...." (3) "Consider that .... This is important because .... Accordingly, we plan to ...."

### Keywords

3-5 keywords that summarize your project.
(e.g., "Keywords: human physiology; bicycle exercise; Gen Z; times-series data")

### Introduction  

Briefly introduce your project.  Include 3-5 research questions. What motivates the questions? Why are they important? (at least 200 words)

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

Assuming you answer your research questions, briefly describe the expected or possible implications for technologists, designers, and policymakers. (at least 150 words)

### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1

Include a chart
```{r, echo = FALSE, code = readLines("Chart 1.R")}
```
Describe why you chose this chart and what information it reveals

### Chart 2

Include a chart
```{r, echo = FALSE, code = readLines("Chart 2.R")}
```

Describe why you chose this chart and what information it reveals

### Chart 3
Include a chart

```{r, echo = FALSE, code = readLines("Chart 3.R")}
```

Describe why you chose this chart and what information it reveals!

In this chart, I have decided to evaluate the Saturation levels between Fenty and Mac Cosmetics in a bar chart model. Saturation is defined as “the level of intensity of a color”. Fenty is represented in Red, while Mac is represented in Blue. As we observe the chart, we can identify that Fenty has a higher saturation level than Mac. Why is that? When we look at the amount of dark shades Fenty produces compared to Mac, Fenty is the overall winner. Darker shades are mainly attributed to how saturated the shade is compared to lighter shades/colors. 


