---
title: "Exploratory Analysis"
output: html_document
---

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

### Authors

Aisha Abdulkarim (aishaa19@uw.edu), Giselle Calvario (gcalva@uw.edu), 
Sumaya Mahamed (sumaha@uw.edu), Aditi Manjunath (aditim4@uw.edu) 

### Date

May 18, 2024

### Abstract

No more than three sentences that summarize your project. Focus on the very most important aspects. For example: (1) "Our main question is .... This question is important because .... To address the question, we will ...." (2) "We are concerned with ..., because .... To address this concern, we plan to ...." (3) "Consider that .... This is important because .... Accordingly, we plan to ...."

### Keywords

Beauty: Foundation shades
Inclusivity: Shade Diversity 
Inclusivity: Possible clientele

### Introduction  

Briefly introduce your project.  Include 3-5 research questions. What motivates the questions? Why are they important? (at least 200 words)

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

This analysis can have many implications. For example, Technologists can explore the underlying mechanisms behind both of the shade-matching algorithms between the brands. By finding out the hue, saturation, and lightness of the foundation shades, it could lead to a better understanding of color-matching algorithms and technology which could lead to more inclusivity as well as accuracy in shades in makeup products in the future. In addition, policymakers can gain insight into the implications of shade inclusivity in the beauty industry for regulations and guidelines. This can help them explore ways to enforce diversity and inclusivity standards in the cosmetic industry and manufacturing of products. This can help ensure that consumers have access to products suiting their diverse skin tones. The interesting implication that the comparison of hue, saturation, and lightness can have is, does the brand has many shades where the shades are pretty similar and not as spread out, or are the shades more spread out and different from each other? This is what we wanted to explore with this analysis. 


### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1: Histogram

```{r, echo = FALSE, code = readLines("Chart 1.R")}
```

Include a chart
Describe why you chose this chart and what information it reveals

### Chart 2: Box Plot

```{r, echo = FALSE, code = readLines("Chart 2.R")}
```

I used a box plot to compare the hues of foundations from Fenty and Mac. From this plot it is evident that is more inclusive of medium and darker skin tones. The range is greater for Fenty foundations hues as well.


### Chart 3: Bar Graph

```{r, echo = FALSE, code = readLines("Chart 3.R")}
```

This Bar Chart is included to show the difference in Saturation Levels of Foundation Shades between Fenty Beauty and Mac Cosmetics. Saturation is defined as “The level of intensity of a color”. The purpose of this graph is to show which brand had the highest saturation level and answer our question of, What role does Saturation levels play in Shade ranges. 

For the bar chart, “Fenty vs. Mac: Saturation Level Differences”, we are evaluating the saturation levels, from the lowest at 0.0 and the highest at 0.8. As we observe the chart, we can identify that Fenty has a higher saturation level than Mac. Why is that? When we look at the number of dark shades Fenty produces compared to Mac, Fenty is the overall winner in producing Dark shades due to its high customer base who have a darker skin tone, while also including the fact that the Company is owned by a woman of color.  Darker shades are mainly attributed to how saturated the shade is compared to lighter shades/colors. When shades have higher results of saturation it results in foundation shades or colors being darker and when the saturation is low the lighter the color, in this case, the lighter the foundation shade. 

According to our dataset, Mac had released 42 Foundation shades, while Fenty had released 39 Foundation shades when this data was taken. Even though Mac seems to have the advantage because they have released more shades compared to Fenty, why is their saturation level low? This might have been influenced by them releasing foundation shades that were lighter and or releasing a smaller variation of dark shades than they have for lighter shades. This is not just something that relates to Mac and Fenty but the majority of beauty companies tend to drop more light-based shades in the early stages of production and then expand to darker shades as they get more consumers. 



### Citations 
cite the sources you used when talking more indept about you charts or about other parts of this project. 

Cosmetics, S. (2023, September 25). How to create and adjust various shades of foundation adapted to skin tones. Sensient Beauty. https://sensient-beauty.com/insights/how-to-create-and-adjust-various-shades-of-foundation-adapted-to-skin-tones/ 